Forecasting the 2024 U.S. Presidential Election: A Poll-of-Polls Approach for Predicting the Outcome

### 2024-11-04 paper/paper.qmd
2

-   **Poll Percentage (pct)**: The `pct` variable is a key predictor as it directly measures the level of support a candidate receives. A higher percentage is expected to lead to a higher probability of winning.
```{r}
#| echo: false

@@

This plot visualizes the relationship between the percentage of poll support (`pct`) and the predicted win probability for each candidate. The red line represents a smoothed curve showing general trends in the data. As expected, higher poll percentages are associated with higher probabilities of winning. This helps confirm that the model captures the influence of poll percentage effectively.

```{r}
#| echo: false
# Plot relationship between poll percentage and predicted win probability
polls_data_cleaned %>%
  ggplot(aes(x = pct, y = predicted_win_prob)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "loess", se = FALSE, color = "red") +
  labs(title = "Relationship Between Poll Percentage and Predicted Win Probability",
       x = "Poll Percentage",
       y = "Predicted Win Probability") +
  theme_minimal()

@@

This boxplot compares the predicted win probabilities for candidates based on their party affiliation (Republican vs. Democratic). It shows that, on average, Democratic candidates have a higher predicted win probability in the polls. This indicates a possible party-related influence on the polling data, as captured by the model.

```{r}
#| echo: false
# Plot predicted win probabilities by party
polls_data_cleaned %>%
  ggplot(aes(x = party_binary, y = predicted_win_prob, fill = factor(party_binary))) +
  geom_boxplot() +
  scale_fill_manual(values = c("0" = "red", "1" = "blue"), labels = c("0" = "Republican", "1" = "Democratic")) +
  labs(title = "Predicted Win Probabilities by Party",
       x = "Party",
       y = "Predicted Win Probability") +
  theme_minimal()

@@

This histogram shows the distribution of predicted win probabilities across all candidates. The distribution reveals how many candidates have higher or lower chances of winning based on polling data. It provides insights into the overall competitiveness of the election, with many candidates clustered around lower probabilities of winning.

```{r}
#| echo: false
# Plot predicted win probability distribution
ggplot(polls_data_cleaned, aes(x = predicted_win_prob)) +
  geom_histogram(binwidth = 0.05, fill = "steelblue", color = "black") +
  labs(title = "Distribution of Predicted Win Probabilities",
       x = "Predicted Win Probability",
       y = "Count") +
  theme_minimal()

>>>>

-   **Poll Percentage (pct)**: The `pct` variable is a key predictor as it directly measures the level of support a candidate receives. A higher percentage is expected to lead to a higher probability of winning.

```{r}
#| echo: false

@@

This plot(@fig-vs) visualizes the relationship between the percentage of poll support (`pct`) and the predicted win probability for each candidate. The red line represents a smoothed curve showing general trends in the data. As expected, higher poll percentages are associated with higher probabilities of winning. This helps confirm that the model captures the influence of poll percentage effectively.

```{r}
#| echo: false
#| label: fig-vs
# Plot relationship between poll percentage and predicted win probability
polls_data_cleaned %>%
  ggplot(aes(x = pct, y = predicted_win_prob)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "loess", se = FALSE, color = "red") +
  labs(title = "Relationship Between Poll Percentage and Predicted Win Probability",
       x = "Poll Percentage",
       y = "Predicted Win Probability") +
  theme_minimal()

@@

This boxplot(@fig-prob) compares the predicted win probabilities for candidates based on their party affiliation (Republican vs. Democratic). It shows that, on average, Democratic candidates have a higher predicted win probability in the polls. This indicates a possible party-related influence on the polling data, as captured by the model.

```{r}
#| echo: false
#| label: fig-prob
# Plot predicted win probabilities by party
polls_data_cleaned %>%
  ggplot(aes(x = party_binary, y = predicted_win_prob, fill = factor(party_binary))) +
  geom_boxplot() +
  scale_fill_manual(values = c("0" = "red", "1" = "blue"), labels = c("0" = "Republican", "1" = "Democratic")) +
  labs(title = "Predicted Win Probabilities by Party",
       x = "Party",
       y = "Predicted Win Probability") +
  theme_minimal()

@@

This histogram(@fig-his) shows the distribution of predicted win probabilities across all candidates. The distribution reveals how many candidates have higher or lower chances of winning based on polling data. It provides insights into the overall competitiveness of the election, with many candidates clustered around lower probabilities of winning.

```{r}
#| echo: false
#| label: fig-his
# Plot predicted win probability distribution
ggplot(polls_data_cleaned, aes(x = predicted_win_prob)) +
  geom_histogram(binwidth = 0.05, fill = "steelblue", color = "black") +
  labs(title = "Distribution of Predicted Win Probabilities",
       x = "Predicted Win Probability",
       y = "Count") +
  theme_minimal()
